---
output:
  html_document:
    keep_md: yes
---

## Introduction

`ADMM` is an R package that utilizes the Alternating Direction Method of Multipliers
(ADMM) algorithm to solve a broad range of statistical optimization problems.
Presently the models that `ADMM` has implemented include Lasso, Elastic Net,
Dantzig Selector, Least Absolute Deviation and Basis Pursuit.

## Models

### Lasso
```{r}
library(glmnet)
library(ADMM)
set.seed(123)
n <- 100
p <- 20
m <- 5
b <- matrix(c(runif(m), rep(0, p - m)))
x <- matrix(rnorm(n * p, mean = 1.2, sd = 2), n, p)
y <- 5 + x %*% b + rnorm(n)

fit <- glmnet(x, y)
out_glmnet <- coef(fit, s = exp(-2), exact = TRUE)
out_admm <- admm_lasso(x, y)$penalty(exp(-2))$fit()
out_paradmm <- admm_lasso(x, y)$penalty(exp(-2))$parallel()$fit()

data.frame(glmnet = as.numeric(out_glmnet),
           admm = as.numeric(out_admm$beta),
           paradmm = as.numeric(out_paradmm$beta))
```

### Elastic Net
```{r}
fit <- glmnet(x, y, alpha = 0.5)
out_glmnet <- coef(fit, s = exp(-2), exact = TRUE)
out_admm <- admm_enet(x, y)$penalty(exp(-2), alpha = 0.5)$fit()
data.frame(glmnet = as.numeric(out_glmnet),
           admm = as.numeric(out_admm$beta))
```

### Dantzig Selector
```{r}
library(flare)
X <- scale(x)
Y <- y - mean(y)

out_flare <- slim(X, Y, nlambda = 20, lambda.min.ratio = 0.01, method = "dantzig")
out_admm <- admm_dantzig(X, Y)$penalty(nlambda = 20, lambda_min_ratio = 0.01)$fit()

range(out_flare$beta - out_admm$beta[-1, ])
```

### Least Absolute Deviation
Least Absolute Deviation (LAD) minimizes `||y - Xb||_1` instead of
`||y - Xb||_2^2` (OLS), and is equivalent to median regression.

```{r}
library(quantreg)
out_rq1 <- rq.fit(x, y)
out_rq2 <- rq.fit(x, y, method = "fn")
out_admm <- admm_lad(x, y, intercept = FALSE)$fit()

data.frame(rq_br = out_rq1$coefficients,
           rq_fn = out_rq2$coefficients,
           admm = out_admm$beta[-1])
```

### Basis Pursuit
```{r}
set.seed(123)
n <- 50
p <- 100
nsig <- 15
beta_true <- c(runif(nsig), rep(0, p - nsig))
beta_true <- sample(beta_true)

x <- matrix(rnorm(n * p), n, p)
y <- drop(x %*% beta_true)
out_admm <- admm_bp(x, y)$fit()

range(beta_true - out_admm$beta)
```


## Performance

### Lasso and Elastic Net

```{r}
library(ADMM)
library(glmnet)
# compute the full solution path, n > p
set.seed(123)
n <- 20000
p <- 1000
m <- 100
b <- matrix(c(runif(m), rep(0, p - m)))
x <- matrix(rnorm(n * p, sd = 2), n, p)
y <- x %*% b + rnorm(n)

system.time(res1 <- glmnet(x, y, nlambda = 20))
system.time(res2 <- admm_lasso(x, y)$penalty(res1$lambda)$fit())
system.time(res3 <- admm_lasso(x, y)$penalty(res1$lambda)$parallel()$fit())

system.time(res4 <- glmnet(x, y, nlambda = 20, alpha = 0.6))
system.time(res5 <- admm_enet(x, y)$penalty(res4$lambda, alpha = 0.6)$fit())

res2$niter
range(coef(res1) - res2$beta)
res3$niter
range(coef(res1) - res3$beta)

res5$niter
range(coef(res4) - res5$beta)

# p > n
set.seed(123)
n <- 2000
p <- 10000
m <- 100
b <- matrix(c(runif(m), rep(0, p - m)))
x <- matrix(rnorm(n * p, sd = 2), n, p)
y <- x %*% b + rnorm(n)

system.time(res1 <- glmnet(x, y, nlambda = 20))
system.time(res2 <- admm_lasso(x, y)$penalty(res1$lambda)$fit())
system.time(res3 <- admm_lasso(x, y)$penalty(res1$lambda)$parallel()$fit())

system.time(res4 <- glmnet(x, y, nlambda = 20, alpha = 0.6))
system.time(res5 <- admm_enet(x, y)$penalty(res4$lambda, alpha = 0.6)$fit())

res2$niter
range(coef(res1) - res2$beta)
res3$niter
range(coef(res1) - res3$beta)

res5$niter
range(coef(res4) - res5$beta)
```

### Dantzig Selector

```{r}
library(ADMM)
library(flare)
# compute the full solution path, n > p
set.seed(123)
n <- 1000
p <- 200
m <- 10
b <- matrix(c(runif(m), rep(0, p - m)))
x <- matrix(rnorm(n * p, sd = 2), n, p)
y <- x %*% b + rnorm(n)

X <- scale(x)
Y <- y - mean(y)

system.time(res1 <- slim(X, Y, nlambda = 20, lambda.min.ratio = 0.01,
                         method = "dantzig"))
system.time(res2 <- admm_dantzig(X, Y)$penalty(nlambda = 20, lambda_min_ratio = 0.01)$
                                      fit())

range(res1$beta - res2$beta[-1, ])
```


### LAD

```{r}
library(ADMM)
library(quantreg)

set.seed(123)
n <- 1000
p <- 500
b <- runif(p)
x <- matrix(rnorm(n * p, sd = 2), n, p)
y <- x %*% b + rnorm(n)

system.time(res1 <- rq.fit(x, y))
system.time(res2 <- rq.fit(x, y, method = "fn"))
system.time(res3 <- admm_lad(x, y, intercept = FALSE)$fit())

range(res1$coefficients - res2$coefficients)
range(res1$coefficients - res3$beta[-1])

set.seed(123)
n <- 5000
p <- 1000
b <- runif(p)
x <- matrix(rnorm(n * p, sd = 2), n, p)
y <- x %*% b + rnorm(n)

system.time(res1 <- rq.fit(x, y, method = "fn"))
system.time(res2 <- admm_lad(x, y, intercept = FALSE)$fit())

range(res1$coefficients - res2$beta[-1])

```
